## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
#if (logical statement) {
  #do something
#} else {
  #do something else
#}
```

The logical statements should boil down to a boolean; for example:

```{r}
a <- 5
if(a>=10){
  print(a)
}else{
  print("it is not more than 10")
}
```

We can also add else if statements in the middle if we need more options.

```{r}
a <- 15
if (a < 5){
  print('small')
}else if (a >=5 & a <=10)  {
 print('medium')
}else if (a >= 10 | a ==7){
  print('large')
}else{
  print('huge')
}
  
```

Neat! We can also nest conditionals just like we do in Python.

```{r}
a <- 5
if(a/2 != 0){
  if(a==5){
    print('a is 5')
  }
}
    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}
for (value in 1:5){
  #paste is a concatenation function
  print(paste(value, "is greater than", value-1))
}
```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}
a_vector <- c(1,1,1,1,1,1,1)
#change the first , third, and sixth elements to the value 6
for(i in c(1,3,6)){
  a_vector[i] <- 6
}

a_vector
```

We can loop over more than just lists:

```{r}
mat <- matrix(1, nrow= 2, ncol = 2)
for(row in 1:2){
  for(col in 1:2){
    mat[row,col] <- 0
    print(mat)
  }
}
```


```{r}
#Like Python, we have while loops in R:
b <- 0
while(b < 10){
  #increase b by 2
  b <- b + 2
  # print b
  print(b)
}
```

```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
  
  
}


```

An example:

```{r}
#start with vector of temperature 
test_temp <- c(16.55678, 26.9420005,32.5910888, 8.580116, 28.003873, 28.5423, 20.976854)                                                                                                 
#Then we want a function that converts from c to f:

c_to_f <- function(data) {
  temp_fahr <- dat * (9/5) + 32
 
  return(temp_fahr)
}


```

print(c_to_f(26))
```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}
data_fahr <- c_to_f(test_temp)
data_fahr
```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}
test_mat <-  matrix(c(1,1,1,2,2,2), nrow <- 2)
test_mat
apply(test_mat,1 , sum) # this adds the value in the vertical or row order, 1 refers to row and 2 is for colomun
apply(test_mat,2 , sum)
```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}
#start with vector of temperature 
test_temp <- c(16.55678, 26.9420005,32.5910888, 8.580116, 28.003873, 28.5423, 20.976854)                                                                                                 
#Then we want a function that converts from c to f:

c_to_f <- function(dat) {
  temp_fahr <- dat * (9/5) + 32
 
  return(temp_fahr)
}

print(c_to_f(26))
```
```{r}
test_temp_mat <- matrix(c(16.55678, 26.9420005,32.5910888, 8.580116, 28.003873, 28.5423, 20.976854, 34.7890), nrow <- 2)

apply(test_temp_mat, 1, c_to_f)
```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:
Q1. Write a function that accepts an integer n and a string s as parameters, and returns a string of s repeated exactly n times.
Examples (input -> output)
6, "I"     -> "IIIIII"
5, "Hello" -> "HelloHelloHelloHelloHello"
    ```{r}
  repeat_string <- function(count, str) {
  if(count > 0){
    return(strrep(str , count));
  }else{
    return('');
  }
}
repeat_string(3, " Ruti")
```
Q2.You take your son to the forest to see the monkeys. You know that there are a certain number there (n), but your son is too young to just appreciate the full number, he has to start counting them from 1.

As a good parent, you will sit and count with him. Given the number (n), populate an array with all numbers up to and including that number, but excluding zero.

For example(Input --> Output):

10 --> [1, 2, 3, 4, 5, 6, 7, 8, 9, 10]
 1 --> [1]

```{r}
monkey_count <- function(n) {
  output <- c()
  for (i in 1:n){
    output <- c(output, i)  
  }
  return(output)
} 
monkey_count((10))
```
```{r}

```
Q3.Write a function that takes an array of numbers and returns the sum of the numbers. The numbers can be negative or non-integer. If the array does not contain any numbers then you should return 0.

Examples
Input: [1, 5.2, 4, 0, -1]
Output: 9.2

Input: []
Output: 0

Input: [-2.398]
Output: -2.398

```{r}
sum_array <- function(a){
  if (length(a) > 0 ){
    return (sum(a))
  }else{
    return (0)
    
  }
}
sum_array(c(1,2,3,5))
```
Q4.We need a function that can transform a number (integer) into a string.

What ways of achieving this do you know?

Examples (input --> output):
123  --> "123"
999  --> "999"
-100 --> "-100"

```{r}
number_to_string <- function(n) {
  return(as.character(n))
}
number_to_string(12345)

```
Q5.We need a function that can transform a string into a number. What ways of achieving this do you know?

Note: Don't worry, all inputs will be strings, and every string is a perfectly valid representation of an integral number.

Examples
"1234" --> 1234
"605"  --> 605
"1405" --> 1405
"-7" --> -7

```{r}
string_to_number <- function(x){
  return(as.integer(x))
}
string_to_number("12345")
```


    ```{r}

```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX




  
```{r}
#You may feel more comfortable figuring this out in Python first, then translating into R later.
  for (i in 0:3) {
    for (j in 0:i) {
      cat("X ")
    }
    cat("\n \n")
  }

```
Create a function that checks an input of type int and returns fizz if the number is a multiple of 3, buzz if the number is a #multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

```{r}
fizz_buzz <- function(n) {
    result <- c()
    for (i in 1:n) {
      if (i %% 3 == 0 ) {
        result <- c(result, "Fizz")
      } else if (i %% 3 == 0 && i %% 5 == 0) {
        result <- c(result, "Fizzbuzz")
      } else if (i %% 5 == 0) {
        result <- c(result, "Buzz")
      } else {
        result <- c(result, as.character(i))
      }
    }
    return(result)
}
fizz_buzz(10)

```